Stream Mining Algorithms for Sensor Data Classification

### 2015-05-27 latex/paper-draft.tex
Yue's changes.

The models present in this paper overcome the disadvantages of incremental neural networks and ensemble classifiers.  Our method of SMNN learn the stream data in mini-batches with sliding windows. In each mini-batch, we run the algorithm in many epochs to reduce the effect of randomized initial weight. Moreover, we pass the weight from one mini-batch to the other mini-batch in order to obtain a better weight. The method of SENN trains multiple neural networks on sliding windows and uses plurality voting on the collection of neural networks to classify new data. Moreover, whenever this is no concept drift, we pass the weights learned from a sliding window to the subsequent sliding window and use them as the initial weights of the next neural network. By doing so, the training  is much faster as compared to using random weights every time.

@@

One of the main requirements for a stream mining algorithm was \emph{constant memory}, which we have not explained in the previous sections yet. This is not an issue for the Sliding Mini-batch Neural Networks algorithm, since the memory used is constant and only depends on the number of weights of the neural network. The same weights matrix is reused when moving from one sliding window to the next, so memory use does not change.

However, it is necessary to consider this issue for Sliding Ensemble Neural Networks, since it involves building a collection of classifiers. Of course, one cannot keep adding models (one per sliding windows) indefinitely.

We fix the maximum number of classifiers that a collection will use as a constant, denoted by $K$. This number is generally much much lower than $n$, the total number of sliding windows.

Models (neural networks) are added to the collection until its size reaches $K$. At this point, the collection is full and we need to determine a way to \emph{continue learning}, so that the algorithm can adapt to the changing data (in the case of concept drift). In order to do so, some models will have to be dropped from the collection so that new models can be included.

We introduce two methods that we designed to pick the models.

@@

	For example, here is a valid way to do this. First, generate a random integer $r$ chosen according to the exponential distribution $\exp(\lambda = 0.1 K)$. Then, let $n = min(floor(r), K)$. We would then discard the model at position $n$ (counting from 0) in our ordered collection of models and replace it with a new model.
\end{enumerate}

@@

\subsection{Results of VFDT and incremental Neural Networks}
need to perform the experiment and obtain the result for comparison.
\subsection{Results of Sliding Mini-Batch Neural Networks}

We looked at the performance of the Sliding Mini-batch Neural Networks as the size of the sliding windows varies.
\begin{figure}[H]
	\centering
	\adjincludegraphics[width=1\linewidth, center]{figures/results-nnet-minibatch}
	\caption{Accuracy of Sliding Mini-Batch Neural Networks classifier on training and testing datasets.}
	\label{fig:benchmark}
\end{figure}
The performance is poor with smaller window sizes since the training sets tend to be poorly balanced. The test accuracy with $L > 20$ is on average 85.5\%. Similarly, the training accuracy for $L > 20$ is on average 98.6\%.

@@

\begin{figure}[H]
	\centering
	\adjincludegraphics[width=1\linewidth, center]{figures/results-nnet-ensemble}
	\caption{Accuracy of ensemble of neural networks on training and testing datasets.}
	\label{fig:results-nnet-ensemble}
\end{figure}

We get 86.0\% test accuracy and 98.5\% training accuracy. This is somewhat surprising for us, as the algorithm appears to overfit, yet the test accuracy is very good.

@@

	%\bibitem{autoencoder_R}
	%Eugene Dubossarsky, et al. Package ‘autoencoder’. CRAN. 2014.

>>>>

		The models present in this paper overcome the disadvantages of incremental neural networks and ensemble classifiers.  Our method of SMNN learn the stream data in mini-batches with sliding windows. In each mini-batch, we run the algorithm in many epochs to reduce the effect of randomized initial weight. Moreover, we pass the weight from one mini-batch to the other mini-batch in order to obtain a better weight. The method of SENN trains multiple neural networks on sliding windows and uses plurality voting on the collection of neural networks to classify new data. Moreover, whenever this is no concept drift, we pass the weights learned from a sliding window to the subsequent sliding window and use them as the initial weights of the next neural network. By doing so, the training  is much faster as compared to using random weights every time. We will also show that the classifiers are diverse enough to give a good prediction.

@@

		The building of this model is counter-intuitive because for an ensemble of classifiers to be more accurate than any of the individual members, it is necessary and sufficient that the classifiers are accurate and diverse \cite{Hansen}. An accurate classifier is one that has an error rate of better than 50\%. Two classifiers are diverse if they don't make errors on the same new data points \cite{Dietterich}. We therefore believe that there is a trade-off between the accuracy and diversity in the ensemble. For our model, by passing the weight when there is no concept drift, we eventually trained the classifier on more data points which increases the accuracy of a single classifier in the ensemble. On the other hand, since the classifiers are built from similar data with related weight, we lose the diversity.

		However, the diversity issue can be solved by the models picking. The models picking means we only pick a constant $K$ classifiers to be in the ensemble. With the principle that we only pick classifiers who are diverse enough to each other, we can have the classifiers in the ensemble more accurate and diverse.

		The models picking also will help to reduce the memory need to run the algorithm. One of the main requirements for a stream mining algorithm was \emph{constant memory}. This is not an issue for the Sliding Mini-batch Neural Networks algorithm, since the memory used is constant and only depends on the number of weights of the neural network. The same weights matrix is reused when moving from one sliding window to the next, so memory use does not change.

		However, it is necessary to consider this issue for Sliding Ensemble Neural Networks, since it involves building a collection of classifiers. Of course, one cannot keep adding models (one per sliding windows) indefinitely. Therefore, we fix the maximum number of classifiers that a collection will use as a constant, denoted by $K$. This number is generally much much lower than $n$, the total number of sliding windows.

		Models (neural networks) are added to the collection until its size reaches $K$. At this point, the collection is full and we need to determine a way to \emph{continue learning}, so that the algorithm can adapt to the changing data (in the case of concept drift). In order to do so, some models will have to be dropped from the collection so that new models can be included. The principle of models picking is that the models in the ensemble are diverse in order to give better prediction.

		We introduce three methods that we designed to pick the models.

@@

			For example, here is a valid way to do this. First, generate a random integer $r$ chosen according to the exponential distribution $\exp(\lambda = 0.1 K)$. Then, let $n = min(floor(r), K)$. We would then discard the model at position $n$ (counting from 0) in our ordered collection of models and replace it with a new model.

			\item The third method is to use error correlation rate to pick up models in order to achieve the diversity in the ensemble. We want to replace those models who give the errors on the same data points. To decide how two models are diverse from each other, we decide to compute the error correlation of two models on a testing dataset. For example, we use all data in one sliding window as the testing data.  compute the e 
		\end{enumerate}

@@

		\subsection{Results of VFDT and incremental Neural Networks}
		For the VFDT, we ran the algorithm hoeffding trees in MOA with the following parameter (after tuning): the grace period= 10, test size=89, and sample Frequency=42 with other parameters as default, this gave us the best performance of VFDT which is 58.43\%. This result is statistically lower than the performance of the batch leaner.
		The reason may be that as the VFDT file reference in VFML \cite{VFML}: ``VFDT  will be most effective when learning from very large data sets, in the millions or billions, where there is plenty of data for it to make good statistical decisions. For smaller data sets you may consider the -batch parameter which makes it load all data into RAM and learn as a traditional decision tree learner. If you suspect that there is concept drift in the data set you may like to use the cvfdt tool instead''.

		On the other hand, the performance of incremental Neural Networks is better. We pass the first 70 days' data one by one into the neural network with updating the weight every time one data instance pass, and the data from the last 89 days as testing data. The only incremental learning achieve a accuracy of 82.14\%.

		\subsection{Results of Sliding Mini-Batch Neural Networks}

		We looked at the performance of the Sliding Mini-batch Neural Networks as the size of the sliding windows varies.
		\begin{figure}[H]
			\centering
			\adjincludegraphics[width=1\linewidth, center]{figures/results-nnet-minibatch}
			\caption{Accuracy of Sliding Mini-Batch Neural Networks classifier on testing datasets.}
			\label{fig:benchmark}
		\end{figure}
		The performance is poor with smaller window sizes since the training sets tend to be poorly balanced. The test accuracy with $L > 20$ is on average 85.5\%.

@@

		\begin{figure}[H]
			\centering
			\adjincludegraphics[width=1\linewidth, center]{figures/results-nnet-ensemble}
			\caption{Accuracy of ensemble of neural networks on  testing datasets.}
			\label{fig:results-nnet-ensemble}
		\end{figure}

		We get 86.0\% test accuracy. This is somewhat surprising for us, as the algorithm appears to overfit, yet the test accuracy is very good.

@@

			\bibitem{VFML}
			Pedro Domingos, Geoff Hulten. "vfdt File Reference". \url{http://www.cs.washington.edu/dm/vfml/vfdt.html}

			\bibitem{Hansen}
			Hansen, Lars Kai, and Peter Salamon. "Neural network ensembles." IEEE transactions on pattern analysis and machine intelligence 12.10 (1990): 993-1001.

			\bibitem{Dietterich}
			Dietterich, Thomas G. "Ensemble methods in machine learning." Multiple classifier systems. Springer Berlin Heidelberg, 2000. 1-15.
			%\bibitem{autoencoder_R}
			%Eugene Dubossarsky, et al. Package ‘autoencoder’. CRAN. 2014.
